# Project 1
# AKSHAY KOCHHAR
# IE6600, 36182 COMPUTATION AND VISUALIZATION, SECTION 2

#Reading all the libraries
```{r}
library(scales)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(tidyr)
library(stringr)
library(usmap)
library(reshape2)
library(readxl)
library(magrittr)
```

#Importing the dataset
```{r}
fmarket <- read.csv("fmarket.csv", na.strings="", stringsAsFactors=FALSE)
View(fmarket)
```

#Fig1: Plotting all the farmers market on US map and showing the highest count
```{r}
fmarket <- read.csv("fmarket.csv", na.strings="", stringsAsFactors=FALSE)
unique(fmarket$State)

farmer_loc <- select(fmarket,city,County,State,Location)
sum(is.na(fmarket$city))
sum(is.na(fmarket$County))
sum(is.na(fmarket$State))
sum(is.na(fmarket$Location))
farmer_loc$State <- as.character(farmer_loc$State)
unique(farmer_loc$State)
# Dividing all the regions according to different states
farmer_loc$Division <- "NA"

farmer_loc$Division <- ifelse(farmer_loc$State %in% c("Connecticut","Maine","Massachusetts","New Hampshire","Rhode Island","Vermont"),"New England",farmer_loc$Division)

farmer_loc$Division <- ifelse(farmer_loc$State %in% c("New Jersey","New York","Pennsylvania"),"Mid-Atlantic",farmer_loc$Division)

farmer_loc$Division <- ifelse(farmer_loc$State %in% c("Virginia", "West Virginia", "Kentucky", "Delaware", "Maryland", "North Carolina", "South Carolina", 
                                                                  "Tennessee", "Arkansas","Louisiana", "Florida", "Georgia", "Alabama","Mississippi","District of Columbia"),"Southern",farmer_loc$Division)

farmer_loc$Division <- ifelse(farmer_loc$State %in% c("Alaska","California", "Oregon","Washington"),"Pacific Coastal",farmer_loc$Division)

farmer_loc$Division <- ifelse(farmer_loc$State %in% c("Texas", "Arizona", "New Mexico","Oklahoma"),"South West", farmer_loc$Division)

farmer_loc$Division <- ifelse(farmer_loc$State %in% c("Montana", "Idaho", "Colorado", "Utah", "Wyoming", "Nevada"),"Rocky Mountain", farmer_loc$Division)

farmer_loc$Division <- ifelse(farmer_loc$State %in% c("Hawaii", "Puerto Rico", "Virgin Islands"),"Islands", farmer_loc$Division)

farmer_loc$Division <- ifelse(farmer_loc$State %in% c("Michigan", "North Dakota", "South Dakota", "Iowa", "Minnesota", "Kansas", "Nebraska", "Ohio", "Indiana", "Illinois", "Wisconsin", "Missouri"),"Mid-west", farmer_loc$Division)

farmer_loc$Division <- as.factor(farmer_loc$Division)
levels(farmer_loc$Division)

temp2 <- plyr::count(farmer_loc, "State")                                                                                
colnames(temp2)[colnames(temp2)=="State"] <- "state"

plot_usmap(data = temp2, values = "freq", color = "red",labels = T,label_color = "black")  + 
  scale_fill_continuous( low = "white", high = "red", name = "Distribution", label = scales::comma ) + 
  theme(legend.position = "right") + labs(title = "US State wise Market Distribution")  + theme(plot.title=element_text(size=14, face="bold",color="mediumorchid4",lineheight=1.4)) + theme(legend.position = c(0.9,0.2)) + theme(legend.title = element_text(colour="Black", size=15, face = "bold" )) + theme(legend.background = element_rect(fill="lightsalmon", size=0.5, linetype="solid"))

```

#Fig2: Count of farmers market based on above divisions
```{r}
fmarket <- read.csv("fmarket.csv", na.strings="", stringsAsFactors=FALSE)
unique(fmarket$State)

farmer_loc <- select(fmarket,city,County,State,Location)
sum(is.na(fmarket$city))
sum(is.na(fmarket$County))
sum(is.na(fmarket$State))
sum(is.na(fmarket$Location))
farmer_loc$State <- as.character(farmer_loc$State)
unique(farmer_loc$State)
farmer_loc$Division <- "NA"

farmer_loc$Division <- ifelse(farmer_loc$State %in% c("Connecticut","Maine","Massachusetts","New Hampshire","Rhode Island","Vermont"),"New England",farmer_loc$Division)

farmer_loc$Division <- ifelse(farmer_loc$State %in% c("New Jersey","New York","Pennsylvania"),"Mid-Atlantic",farmer_loc$Division)

farmer_loc$Division <- ifelse(farmer_loc$State %in% c("Virginia", "West Virginia", "Kentucky", "Delaware", "Maryland", "North Carolina", "South Carolina", 
                                                                  "Tennessee", "Arkansas","Louisiana", "Florida", "Georgia", "Alabama","Mississippi","District of Columbia"),"Southern",farmer_loc$Division)

farmer_loc$Division <- ifelse(farmer_loc$State %in% c("Alaska","California", "Oregon","Washington"),"Pacific Coastal",farmer_loc$Division)

farmer_loc$Division <- ifelse(farmer_loc$State %in% c("Texas", "Arizona", "New Mexico","Oklahoma"),"South West", farmer_loc$Division)

farmer_loc$Division <- ifelse(farmer_loc$State %in% c("Montana", "Idaho", "Colorado", "Utah", "Wyoming", "Nevada"),"Rocky Mountain", farmer_loc$Division)

farmer_loc$Division <- ifelse(farmer_loc$State %in% c("Hawaii", "Puerto Rico", "Virgin Islands"),"Islands", farmer_loc$Division)

farmer_loc$Division <- ifelse(farmer_loc$State %in% c("Michigan", "North Dakota", "South Dakota", "Iowa", "Minnesota", "Kansas", "Nebraska", "Ohio", "Indiana", "Illinois", "Wisconsin", "Missouri"),"Mid-west", farmer_loc$Division)

farmer_loc$Division <- as.factor(farmer_loc$Division)
levels(farmer_loc$Division)

temp1 <- plyr::count(farmer_loc, "Division")

ggplot(temp1, aes(x= reorder(factor(Division),freq),y=freq, fill = Division)) + geom_bar(stat="identity") + coord_flip() +
  geom_text(aes(label=freq),hjust = 1.1, size=3.2) + theme_minimal()  + labs(title = "Distribution of Farmers Market based on Region") + xlab("Region") + ylab("Count of Farmer Markets") +  theme(plot.title=element_text(size=14, face="bold",color="Purple",lineheight=1.3)) + theme(axis.title.x = element_text(color="Purple", size=14, face="bold"),
                                                                                                                                                                                                    axis.title.y = element_text(color="Purple", size=14, face="bold"))

```

#Fig3: Trend of farmer markets
```{r}
temp3 <- as.data.frame(fmarket$Season1Date)
colnames(temp3)[1] <- "Season1Date"
#separating the year and getting just the date
temp3 <- as.data.frame(temp3[grepl("\\/", temp3[["Season1Date"]])==TRUE,])
colnames(temp3) [1]<- "Season1Date"
temp3$year1 <- substr(temp3$Season1Date, 7, 10)
temp3 <- as.data.frame(temp3[grepl("Date|\\/", temp3[["year1"]])==FALSE,])
temp3.1 <- plyr::count(temp3$year1)
ggplot(temp3.1, aes(x = x, freq, group = 1)) +
  geom_line(aes(color = x), stat = "identity") +
  geom_text(aes(label = freq), vjust = -0.5, size=3.6) +
  ggtitle("Year wise Farmers Market (Season 1)") + 
  xlab("Year") +
  ylab("Count of Farmers Market") +
  theme(text=element_text(size=14,  family="Arial"), legend.position = 'none',
        plot.title=element_text(hjust = 0.6)) + theme(plot.title=element_text(size=14, face="bold",color="Violetred",lineheight=1.3)) +theme(axis.title.x = element_text(color="Violetred", size=14, face="bold"),
                                                      axis.title.y = element_text(color="Violetred", size=14, face="bold"))
  geom_point(aes(color = x))

```

#Fig4: Different Payment methods
```{r}
#Selecting different payment methods from the dataset
temp5 <- fmarket[,c(1,24,25,26,27,28)]
temp5.1 <- as.data.frame(ifelse(temp5[,2:6] == "Y","1","0"))
apply(is.na(temp5.1),2,sum)
temp5.2 <- as.data.frame(lapply(temp5.1,as.numeric))
temp5.2 <- as.data.frame(ifelse(temp5.2 == 2,1,0))

sum_data=data.frame(value=apply(temp5.2,2,sum))
sum_data$key=rownames(sum_data)

ggplot(data=sum_data, aes(x=reorder(key,value), y=value, fill=key)) +
  geom_bar(stat="identity") +  geom_text(aes(label=value),hjust = 0.6, size=6, vjust = 2) + theme_light()  + labs(title = "Payment Methods") + xlab("Payment Type") + ylab("Count of Farmer Markets") +theme(legend.position = "none") + theme(plot.title=element_text(size=19, face="bold",color="navyblue",lineheight=1.3, hjust = 0.4)) + theme(axis.title.x = element_text(color="navyblue", size=14, face="bold"),
                                                                                                                                                                                                                                                                                                                                                   axis.title.y = element_text(color="navyblue", size=14, face="bold"))


```

#Fig5: Top SNAP Providing states
```{r}
X1<-fmarket %>%
  select(State,Credit,WIC,WICcash,SFMNP,SNAP)%>%
  na.omit()
y1 <- melt(X1, id.vars = "State")

T1<-y1 %>%
  filter(variable=="SNAP")%>%
  filter(value=="Y")%>%
  group_by(State)%>%
  summarise(Count=n())

U1 <- T1 %>% top_n(10)
ggplot(U1, aes(x=reorder(State,-Count), y=Count)) + 
  geom_bar(stat = "identity",fill="#E69F00") +geom_text(aes(label=Count),hjust = 0.7, vjust = 1.7, size=3.2) +
  theme_minimal()+theme(axis.text.x = element_text(angle = 90, vjust=0.6, hjust=1))+xlab("States")+
  ylab("Count")+ggtitle("Top 10 SNAP providing States")+theme(legend.position = "none")+theme(plot.title=element_text(size=16, face="bold",color="Black",lineheight=1.2))

```

#Fig6: Top SFMNP providing states
```{r}
X1<-fmarket %>%
  select(State,Credit,WIC,WICcash,SFMNP,SNAP)%>%
  na.omit()
y1 <- melt(X1, id.vars = "State")

A1<-y1 %>%
  filter(variable=="SFMNP")%>%
  filter(value=="Y")%>%
  group_by(State)%>%
  summarise(Count=n())

M1 <- A1 %>% top_n(10)
ggplot(M1, aes(x=reorder(State,-Count), y=Count)) + 
  geom_bar(stat = "identity",fill="#D55E00") +geom_text(aes(label=Count),hjust = 0.7, vjust = 1.7, size=3.2) +
  theme_minimal()+theme(axis.text.x = element_text(angle = 90, vjust=0.6, hjust=1))+xlab("States")+
  ylab("Count")+ggtitle("Top 10 SFMNP providing States")+theme(legend.position = "none")+theme(plot.title=element_text(size=16, face="bold",color="Black",lineheight=1.2))
```

#Fig7: Top WIC providing states
```{r}
X1<-fmarket %>%
  select(State,Credit,WIC,WICcash,SFMNP,SNAP)%>%
  na.omit()
y1 <- melt(X1, id.vars = "State")

L1<-y1 %>%
  filter(variable=="WIC")%>%
  filter(value=="Y")%>%
  group_by(State)%>%
  summarise(Count=n())

U1 <- L1 %>% top_n(10)
ggplot(U1, aes(x=reorder(State,-Count), y=Count)) + 
  geom_bar(stat = "identity",fill="#FF9933") +geom_text(aes(label=Count),hjust = 0.7, vjust = 1.7, size=3.2) +
  theme_minimal()+theme(axis.text.x = element_text(angle = 90, vjust=0.6, hjust=1))+xlab("States")+
  ylab("Count")+ggtitle("Top 10 WIC providing States")+theme(legend.position = "none")+theme(plot.title=element_text(size=16, face="bold",color="Black",lineheight=1.2))

```

#Fig8: Top WICcash accepting states
```{r}
X1<-fmarket %>%
  select(State,Credit,WIC,WICcash,SFMNP,SNAP)%>%
  na.omit()
y1 <- melt(X1, id.vars = "State")

H1<-y1 %>%
  filter(variable=="WICcash")%>%
  filter(value=="Y")%>%
  group_by(State)%>%
  summarise(Count=n())

F1 <- H1 %>% top_n(10)
ggplot(F1, aes(x=reorder(State,-Count), y=Count)) + 
  geom_bar(stat = "identity",fill="#33FF99") +geom_text(aes(label=Count),hjust = 0.7, vjust = 1.7, size=3.2) +
  theme_minimal()+theme(axis.text.x = element_text(angle = 90, vjust=0.6, hjust=1))+xlab("States")+
  ylab("Count")+ggtitle("Top 10 WICcash providing States")+theme(legend.position = "none")+theme(plot.title=element_text(size=16, face="bold",color="Black",lineheight=1.2))
```

#Fig9: Count of all farmer markets producing different items
```{r}
temp7.1 <- fmarket[,29:58]
temp7.2 <- as.data.frame(ifelse(temp7.1[,2:30] == "Y","1","0"))
temp7.3 <- na.omit(temp7.2, cols = "Bakedgoods")
apply(is.na(temp7.3),2,sum)
temp7.3 <- as.data.frame(lapply(temp7.3,as.numeric))
temp7.3 <- as.data.frame(ifelse(temp7.3 == 2,1,0))
sum_data1=data.frame(value=apply(temp7.3,2,sum))
sum_data1$key=rownames(sum_data1)
allproducts12<- sumdata[order(-sum_data1$value),][1:29,]
ggplot(data=allproducts12, aes(x=reorder(key,value), y=value, fill=key)) +
  geom_bar(position="Stack", colour="black", stat="identity") + coord_flip() +  geom_text(aes(label=value),hjust = 1.5, size=3.2) + theme_minimal()  + labs(title = "All Products availability") + xlab("Product") + ylab("Number of Farmer Markets") +theme(legend.position = "none") + theme(plot.title=element_text(size=18, face="bold",color="sienna",lineheight=1.2, hjust = 0.4))  + theme(axis.title.x = element_text(color="mediumvioletred", size=14, face="bold"),
                                                                                                                                                                                                                                                                                                                                                                                                     axis.title.y = element_text(color="mediumvioletred", size=14, face="bold"))

```

#Fig10: All products available (Grouped)
```{r}
temp7.4 <- temp7.3
#Grouping of all the products based on below mentioned groups
temp7.4$Ready_to_eat <- rowSums(temp7.3[,c(1,17)])
temp7.4$Non_Edible <- rowSums(temp7.3[,c(3,18)])
temp7.4$Plant_Based <- rowSums(temp7.3[,c(4,7,8,14,15,19,22,23,24)])
temp7.4$Meat_Based <- rowSums(temp7.3[,c(6,12,16)])
temp7.4$Drinks <- rowSums(temp7.3[,c(20,21,25)])
temp7.4$Spreads <- rowSums(temp7.3[,c(9,10,11)])
temp7.5 <- temp7.4[,c(2,5,13,26,27,28,29,30,31,32,33,34)]
sum.data12=data.frame(value=apply(temp7.5,2,sum))
sum.data12$key=rownames(sum.data12)
ggplot(data=sum.data12, aes(x=reorder(key,value), y=value, fill=key)) +
  geom_bar(colour="black", stat="identity") + coord_flip() +  geom_text(aes(label=value), hjust = 0.01, size=3.2) + theme_minimal()  + labs(title = "Product Avalability(Groups)") + xlab("Product") + ylab("Number of Market") +theme(legend.position = "none") + theme(plot.title=element_text(size=18, face="bold",color="navyblue",lineheight=1.2, hjust = 0.4))  + theme(axis.title.x = element_text(color="navyblue", size=14, face="bold"),
                                                                                                                                                                                                                                                                                                                                                                     axis.title.y = element_text(color="navyblue", size=14, face="bold"))

```

#Fig11: Top 10 coffee producing states
```{r}
X2<-fmarket %>%
  select(State,Coffee, Honey, Poultry)%>%
  na.omit()
y2 <- melt(X2, id.vars = "State")

H2<-y2 %>%
  filter(variable=="Coffee")%>%
  filter(value=="Y")%>%
  group_by(State)%>%
  summarise(Count=n())

F2 <- H2 %>% top_n(10)
ggplot(F2, aes(x=reorder(State,-Count), y=Count)) + 
  geom_bar(stat = "identity",fill="Lightblue") +geom_text(aes(label=Count),hjust = 0.7, vjust = 1.7, size=3.2) +
  theme_minimal()+theme(axis.text.x = element_text(angle = 90, vjust=0.6, hjust=1))+xlab("States")+
  ylab("Count")+ggtitle("Top 10 Coffee producing States")+theme(legend.position = "none")+theme(plot.title=element_text(size=16, face="bold",color="Black",lineheight=1.2))

```

